---
title: "Ambiguity and Anaphora - An Analysis of Disagreement in Anaphoric Annotation"
author: "Lukas Beiske"
date: "29 10 2024"
output: html_document
---

## Libraries

```{r message = FALSE, echo=TRUE, warning=FALSE, results="hide"}
library(xml2)
library(dplyr)
library(readr)
library(ggplot2)
library(knitr)
library(stringr)
library(tidyr)
library(kableExtra)
library(grid)
library(gridExtra)
```

## Data Treatment

Since information about markables is stored in XML files, these are first converted into CSV format for better handling. The following code is a function that is responsible for this conversion. It loads an XML file, takes care of the XML namespaces, and then extracts the information of the individual markables. A for-loop makes it possible to repeat this process for all XML files in the folder. The output is stored in a separate folder called 'output_csv' within the input folder and consists of one CSV file per XML input file.

```{r xml to csv, message = FALSE, echo=TRUE, warning=FALSE}
#define the function; argument = individual XML files instantiated below
xml_to_csv <- function(xml_ind, output_f) {
  #read the file via read_xml (xml2)
  xml_data <- read_xml(xml_ind)
  #strip default namespaces via xml_ns_stripfrom document (xml2)
  xml_data <- xml_ns_strip(xml_data)
  #data extraction
  xml_data %>%
    #retrieve all attributes values as a named character vector, bind rows and create csv-file (name derived from original XML file)
    xml_find_all("markable") %>% 
    xml_attrs() %>%
    bind_rows() %>%
    write_csv(file = file.path(output_f, paste0(sub(".xml", ".csv", basename(xml_ind)))))
}

#path to input folders with XML files
folders_xml <- c("../Statistics/data/ARRAU/RST",
                 "../Statistics/data/ARRAU/Trains_91",
                 "../Statistics/data/PD/gutenberg",
                 "../Statistics/data/PD/wiki")

#for-loop for iterating through input folders
for (input_f in folders_xml) {
  #create a list of contained XML files
  all_xml_ind <- list.files(path = input_f, pattern = "\\.xml$", full.names = TRUE)
  #create output folders
  output_f <- file.path(input_f, "output_csv")
  dir.create(output_f, showWarnings = FALSE)
  
  for (xml_ind in all_xml_ind) {
    xml_to_csv(xml_ind, output_f)
  }
}
```

### Create a DF for ARRAU data

The following code creates data frames from the data of the ARRAU corpus.

```{r message=FALSE, results='hide'}

#define function
create_df <- function(path_f, name_ds) {
  #create a list with all CSV files
  csv_list <- list.files(path_f, pattern = "\\.csv$", full.names = TRUE)
  #create empty data frame 
  dataset <- data.frame()
  #for-loop to iterate through the CSV files and extract the data
  for (csv_ind in csv_list) {
    csv_data <- read_csv(csv_ind)
    #add the extra columns for data-set and file name
    csv_data$dataset <- name_ds
    csv_data$file_name <- basename(csv_ind)
    #combine the data of the individual files
    dataset <- bind_rows(dataset, csv_data)
  }
  return(dataset)
}

#define the paths to the folders containinng the CSV files
RST_path <- "../Statistics/data/ARRAU/RST/output_csv"
TRAINS91_path <- "../Statistics/data/ARRAU/Trains_91/output_csv"

#apply function to the CSV files of the RST domain; create df
RST_df <- create_df(RST_path, "RST")

#apply function to the CSV files of the TRAINS domain; create df
TRAINS91_df <- create_df(TRAINS91_path, "Trains_91")

#combine the RST_df and TRAINS91_df data frames
COMB_RST_TRAINS_df <- bind_rows(RST_df, TRAINS91_df)

```

### Create a DF for Phrase Detectives (PD) data

The following code creates data frames from the data of the PD corpus.

```{r message=FALSE, results='hide'}

#define function
create_df <- function(path_f, name_ds) {
  #create a list with all CSV files
  csv_list <- list.files(path_f, pattern = "\\.csv$", full.names = TRUE)
  #create empty data frame 
  dataset <- data.frame()
  #for-loop to iterate through the CSV files and extract the data
  for (csv_ind in csv_list) {
    csv_data <- read_csv(csv_ind)
    #add the extra columns for data-set and file name
    csv_data$dataset <- name_ds
    csv_data$file_name <- basename(csv_ind)
    #combine the data of the individual files
    dataset <- bind_rows(dataset, csv_data)
  }
  return(dataset)
}

#define the paths to the folders containinng the CSV files
GB_path <- "../Statistics/data/PD/gutenberg/output_csv"
WIKI_path <- "../Statistics/data/PD/wiki/output_csv"

#apply function to the CSV files of the gutenberg domain; create df
GB_df <- create_df(GB_path, "GB")

#apply function to the CSV files of the wikipedia domain; create df
WIKI_df <- create_df(WIKI_path, "Wiki")

#combine the RST_df and TRAINS91_df data frames
COMB_GB_WIKI_df <- bind_rows(GB_df, WIKI_df)

```

```{r}
## combine ARRAU and PD data; removal of columns so that both observe the same

# Remove 'disagreement_type' column from COMB_GB_WIKI_df; new df called CGW
CGW <- COMB_GB_WIKI_df[, -which(names(COMB_GB_WIKI_df) == "disagreement_type"), drop = FALSE]

# Remove multiple columns from COMB_RST_TRAINS_df; new df is called CRT
CRT <- COMB_RST_TRAINS_df[, -which(names(COMB_RST_TRAINS_df) %in% c("on_map", "on_map_2", "objectadditional")), drop = FALSE]

# Combine CRT and CGW in a new df called combined_df
combined_df <- rbind(CRT, CGW)#FOR FURTHER ANALYSIS

##sortcolumns
#general information in fixed order
general_info <- c("min_words", "min_ids", "id", "file_name", "span", "mmax_level", "dataset", "comment")

#further (gramm.) information in fixed order
info_gramm <- c("number", "gender", "person", "gram_fnc", "ambiguity")

#annotated information sorted alphabetically
all_columns <- names(combined_df)
anno_info <- setdiff(all_columns, c(general_info, info_gramm))  # Find remaining columns
anno_info_sorted <- sort(anno_info)  # Sort alphabetically

# Reorder DataFrame by combining these three groups
df_sorted <- combined_df[, c(general_info, info_gramm, anno_info_sorted)]

# Make all character via list, then put it back in a data frame
df_sorted <- lapply(df_sorted, as.character)

df_sorted <- as.data.frame(df_sorted)

# View the reordered DataFrame
#print(df_sorted)
```


## Categorisation

```{r}
#filter for only those cases that are annotated as being ambiguous
combined_df_amb <- df_sorted %>%
  filter(ambiguity == "ambiguous")
```

```{r}
###instantiate auxiliary columns
#for analysis
combined_df_amb$info_status_1 <- NA #info status of 1st reading
combined_df_amb$info_status_2 <- NA #info status of 2nd reading
combined_df_amb$amb_generic <- NA #ambiguity in generic value 
combined_df_amb$amb_category <- NA #ambiguity in category value
combined_df_amb$amb_bridging <- NA #ambiguity regarding a bridging relationship

#for concatenation of the final label
combined_df_amb$amb_reason_pre <- NA #label for ambiguity regarding info status, bridging, etc. to concatenate
combined_df_amb$amb_reason_gen <- NA #label for ambiguity in generic value to concatinate
combined_df_amb$amb_reason_cat <- NA #label for ambiguity in category value to concatinate
combined_df_amb$amb_reason_fin <- NA #result of concatenating the three labels above
```

##### Fill in Auxiliary Columns

```{r}
#info_status_1
combined_df_amb$info_status_1 <- ifelse(
  combined_df_amb$reference == "old",
  "DO",
  ifelse(
    combined_df_amb$reference == "new",
    "DN",
    ifelse(
      combined_df_amb$reference == "non_referring",
      "non_referring",
      NA  # Assign NA for other cases
    )
  )
)

#info_status_2
combined_df_amb$info_status_2 <- ifelse(
  combined_df_amb$ref_type_2 %in% c("phrase", "segment"),
  "DO",
  ifelse(
    combined_df_amb$ref_type_2 == "new",
    "DN",
    ifelse(
      combined_df_amb$ref_type_2 == "non_referring",
      "non_referring",
      ifelse(
        combined_df_amb$ref_type_2 == "undef_reference",
        "undef_reference",
        ifelse(
          combined_df_amb$ref_type_2 == "undef_reference_2",
          "undef_reference_2",
          NA  # Assign NA for other cases
        )
      )
    )
  )
)

#amb_generic
###checked: when NA was assigned for generic/generic_2 -> amb_generic = NA (51), ref_type_2 = segment (51). This means, if there is an NA in generic_2 now, "no" is assigned, since this only happens when ref_typ_2 = segment, in which case generic_2 cannot be assigned according to the scheme, thus, an ambiguity between the generic values cannot be evaluated
combined_df_amb$amb_generic <- ifelse(
  is.na(combined_df_amb$generic_2) | is.na(combined_df_amb$generic),
  "no", #only when ref_type_2 = segment -> generic_2 = NA
  ifelse(
    combined_df_amb$generic != combined_df_amb$generic_2,
    "yes",
    "no"
        )
  )

#amb_category 
combined_df_amb$amb_category <- ifelse(
  is.na(combined_df_amb$category) | is.na(combined_df_amb$category_2),
  "no",
  ifelse(
    combined_df_amb$category != combined_df_amb$category_2,
    "yes",
    "no"
        )
  )

#amb_bridging
combined_df_amb$amb_bridging <- ifelse(
  is.na(combined_df_amb$related_object) | is.na(combined_df_amb$related_object_2),
  "no", 
  ifelse(
    xor(combined_df_amb$related_object == "yes", combined_df_amb$related_object_2 == "yes"),
    "yes",
    "no"
  )
)
```


```{r}
#fill auxiliary columns for concatinating the final label
#amb_reason_gen
combined_df_amb$amb_reason_gen <- case_when(
  combined_df_amb$amb_generic == "yes" ~ "generic-value",
  combined_df_amb$amb_generic == "no" ~ "",
  TRUE ~ "CHECK_gen" #to verify if the assignment works correctly
)

#amb_reason_cat
combined_df_amb$amb_reason_cat <- case_when(
  combined_df_amb$amb_category == "yes" ~ "category",
  combined_df_amb$amb_category == "no" ~ "",
  TRUE ~ "CHECK_cat" #to verify if the assignment works correctly
)
```

##### Referring vs. Non-referring

```{r}
###referring vs. non-referring
#df_non_ref
df_non_ref <- combined_df_amb %>%
  filter(reference == "non_referring" | ref_type_2 == "non_referring") #the content of df_non_ref is always ambiguous between referring and non_referring

#assign label
df_non_ref$amb_reason_pre <- "referring vs. non_referring" #this can be done, because of the filter above
```

##### Referring vs. Undef_reference

```{r}
###referring vs. undefined reference
#df_ref_undef_ref
df_ref_undef_ref <- combined_df_amb %>%
  filter(ref_type_2 == "undef_reference" | ref_type_2 == "undef_reference_2") #when reference = non_referring or undef_reference, no annotation as ambiguous possible

#assign label
df_ref_undef_ref$amb_reason_pre <- "referring vs. undefined_reference"
```

##### Referring vs. Referring

```{r}
# filter only for referring mentions and exclude potential unmarked markables
combined_df_ref <- combined_df_amb %>%
  filter(!(reference %in% c("unmarked", "non_referring")) & ref_type_2 != "non_referring") 
```

##### DN vs. DN

```{r}
###DN_DN
#categorisation using filters
#checked count of original df (df_DN_DN) and resulting df (df_DN_DN_fin) is the same

#filter for cases that are new in both readings
df_DN_DN <- combined_df_ref %>%
  filter(info_status_1 == "DN" & info_status_2 == "DN")

#this filter is used to identify DN vs. DN cases. Later they will get assigned an amb for generic and/or category
#all markables are amb in either generic or category or both except one -> second reading has an object_2 
df_DN_DN_only <- df_DN_DN %>%
  filter(related_object == "no" & related_object_2 == "no")
#assign label
df_DN_DN_only$amb_reason_pre <- "DN vs. DN, no related objects"

#this filter is used to identify DN vs. DN+bridging cases
df_DN_DN_a <- df_DN_DN %>%
  filter(amb_bridging == "yes")
#assign label
df_DN_DN_a$amb_reason_pre <- "DN vs. DN+bridging"

#this filter is used to identify cases where the right entity for bridging was not identified but the relation is the same
df_DN_DN_b <- df_DN_DN %>%
  filter(related_object == "yes" & related_object_2 == "yes" & related_phrase != related_phrase_2 & related_rel == related_rel_2)
#asssign label
df_DN_DN_b$amb_reason_pre <- "identifying the right antecedent for bridging (DN)"

#this filter is used to identify cases where the right entity for bridging was not identified and the relation is not the same
df_DN_DN_c <- df_DN_DN %>%
  filter(related_object == "yes" & related_object_2 == "yes" & related_phrase != related_phrase_2 & related_rel != related_rel_2)
#asssign label
df_DN_DN_c$amb_reason_pre <- "identifying the right antecedent for bridging (DN) and ambiguous relation"

#this filter is used to identify cases where the bridging relation is ambigous
df_DN_DN_d <- df_DN_DN %>%
  filter(related_object == "yes" & related_object_2 == "yes" & related_phrase == related_phrase_2 & related_rel != related_rel_2)
#asssign label
df_DN_DN_d$amb_reason_pre <- "DN vs. DN; ambiguous relation"

#this filter is used for cases where neither the related phrase nor the related relation is ambiguous. However there could be cases where category or generic is amb (in fact, only one)
df_DN_DN_e <- df_DN_DN %>%
  filter(related_object == "yes" & related_object_2 == "yes" & related_phrase == related_phrase_2 & related_rel == related_rel_2)
#asssign label
df_DN_DN_e$amb_reason_pre <- "DN vs. DN"


#recombine the dfs
df_DN_DN_fin <- rbind(df_DN_DN_only, df_DN_DN_a, df_DN_DN_b, df_DN_DN_c, df_DN_DN_d, df_DN_DN_e)

#### -----> exclusion of one observation: DN vs. DN, no related object. It is amb in on_map, which is not evaluated here.

df_DN_DN_fin <- df_DN_DN_fin %>%
  filter(!(min_ids == "word_69" & file_name == "dia6-1_phrase_level.csv" & id == "markable_17"))


#print(df_DN_DN_fin)

```

##### DN vs. DO

```{r}
###DN_DO
df_DN_DO <- combined_df_ref %>%
  filter(info_status_1 == "DN" & info_status_2 == "DO")

#this filter is used to identy cases that are DO vs. DN without further attributes; single phrase antecedent
df_DN_DO_a <- df_DN_DO %>%
  filter(reference == "new" & ref_type_2 == "phrase" &
           phrase_antecedent_2 == "single_phrase" &
           related_object == "no" & related_object_2 == "no"
           )
#assign label
df_DN_DO_a$amb_reason_pre <- "DO (single) vs. DN"

#this filter is used to identy cases that are DO+bridging vs. DN; single phrase antecedent --------------------------> no such cases
# df_DN_DO_a_1 <- df_DN_DO %>%
#   filter(reference == "new" & ref_type_2 == "phrase" &
#            phrase_antecedent_2 == "single_phrase" &
#            related_object == "no" & related_object_2 == "yes"
#            )
# #assign label
# df_DN_DO_a_1$amb_reason_pre <- "DO+bridging (single) vs. DN"

#this filter is used to identy cases that are DO vs. DN+bridging single phrase antecedent 
df_DN_DO_a_2 <- df_DN_DO %>%
  filter(reference == "new" & ref_type_2 == "phrase" &
           phrase_antecedent_2 == "single_phrase" &
           related_object == "yes" & related_object_2 == "no"
           )
#assign label
df_DN_DO_a_2$amb_reason_pre <- "DO (single) vs. DN+bridging"

#this filter is used to identy cases that are DO (singel) + bridging vs. DN+bridging single phrase antecedent; same relation --------------------------> no such cases
# df_DN_DO_a_3 <- df_DN_DO %>%
#   filter(reference == "new" & ref_type_2 == "phrase" &
#            phrase_antecedent_2 == "single_phrase" &
#            related_object == "yes" & related_object_2 == "yes" &
#            related_rel == related_rel_2
#            )
# #assign label
# df_DN_DO_a_3$amb_reason_pre <- "DO+bridging (single) vs. DN+bridging"

#this filter is used to identy cases that are DO (singel) + bridging vs. DN+bridging single phrase antecedent; different relation --------------------------> no such cases
# df_DN_DO_a_4 <- df_DN_DO %>%
#   filter(reference == "new" & ref_type_2 == "phrase" &
#            phrase_antecedent_2 == "single_phrase" &
#            related_object == "yes" & related_object_2 == "yes" &
#            related_rel != related_rel_2
#            )
# #assign label
# df_DN_DO_a_4$amb_reason_pre <- "DO+bridging (single) vs. DN+bridging; ambiguous relation"

#this filter is used to identy cases that are DO vs. DN without further attributes; multiple phrase antecedents --------------------------> no such cases
# df_DN_DO_b <- df_DN_DO %>%
#   filter(reference == "new" & ref_type_2 == "phrase" &
#            phrase_antecedent_2 == "multiple_phrases" &
#            related_object == "no" & related_object_2 == "no"
#            )
# #assign label
# df_DN_DO_b$amb_reason_pre <- "DO (multiple) vs. DN"

#this filter is used to identy cases that are DO+bridging vs. DN ; multiple phrase antecedents --------------------------> no such cases
# df_DN_DO_b_1 <- df_DN_DO %>%
#   filter(reference == "new" & ref_type_2 == "phrase" &
#            phrase_antecedent_2 == "multiple_phrases" &
#            related_object == "no" & related_object_2 == "yes"
#            )
# #assign label
# df_DN_DO_b_1$amb_reason_pre <- "DO+bridging (multiple) vs. DN"

#this filter is used to identy cases that are DO vs. DN+bridging ; multiple phrase antecedents 
df_DN_DO_b_2 <- df_DN_DO %>%
  filter(reference == "new" & ref_type_2 == "phrase" &
           phrase_antecedent_2 == "multiple_phrases" &
           related_object == "yes" & related_object_2 == "no"
           )
#assign label
df_DN_DO_b_2$amb_reason_pre <- "DO (multiple) vs. DN+bridging"

# #this filter is used to identy cases that are DO+bridging vs. DN+bridging ; multiple phrase antecedents, related relation is the same --------------------------> no such cases
# df_DN_DO_b_3 <- df_DN_DO %>%
#   filter(reference == "new" & ref_type_2 == "phrase" &
#            phrase_antecedent == "multiple_phrases" &
#            related_object == "yes" & related_object_2 == "yes" &
#            related_rel == related_rel_2
#            )
#assign label
# df_DN_DO_b_3$amb_reason_pre <- "DO+bridging (multiple) vs. DN+bridging"

#this filter is used to identy cases that are DO+bridging vs. DN+bridging ; multiple phrase antecedents, related relation is not the same --------------------------> no such cases
# df_DN_DO_b_4 <- df_DN_DO %>%
#   filter(reference == "new" & ref_type_2 == "phrase" &
#            phrase_antecedent_2 == "multiple_phrases" &
#            related_object == "yes" & related_object_2 == "yes" &
#            related_rel != related_rel_2
#            )
# #assign label
# df_DN_DO_b_4$amb_reason_pre <- "DO+bridging (multiple) vs. DN+bridging; ambiguous related relation"

#this filter is used to identy cases that are DO+deixis vs. DN without further attributes
df_DN_DO_c <- df_DN_DO %>%
  filter(reference == "new" & ref_type_2 == "segment" &
           related_object == "no" & related_object_2 == "no"
           )
#assign label
df_DN_DO_c$amb_reason_pre <- "deixis vs. DN"

#this filter is used to identy cases that are DO+deixis+bridging vs. DN without further attributes --------------------------> no such cases
# df_DN_DO_c_1 <- df_DN_DO %>%
#   filter(reference == "new" & ref_type_2 == "segment" &
#            related_object == "no" & related_object_2 == "yes"
#            )
# #assign label
# df_DN_DO_c_1$amb_reason_pre <- "deixis+bridging vs. DN"

#this filter is used to identy cases that are DO+deixis vs. DN+bidging without further attributes
df_DN_DO_c_2 <- df_DN_DO %>%
  filter(reference == "new" & ref_type_2 == "segment" &
           related_object == "yes" & related_object_2 == "no"
           )
#assign label
df_DN_DO_c_2$amb_reason_pre <- "deixis vs. DN+bridging"

#this filter is used to identy cases that are DO+deixis+bridging vs. DN+bidging without further attributes --------------------------> no such cases
# df_DN_DO_c_3 <- df_DN_DO %>%
#   filter(reference == "new" & ref_type_2 == "segment" &
#            related_object == "yes" & related_object_2 == "yes"
#            )
# #assign label
# df_DN_DO_c_3$amb_reason_pre <- "deixis+briidging vs. DN+bridging"


# recombine the dfs
df_DN_DO_fin <- rbind(df_DN_DO_a, df_DN_DO_a_2, df_DN_DO_b_2, df_DN_DO_c, df_DN_DO_c_2)

#print(df_DN_DO_fin)
```

##### DO vs. DN

```{r}
###DO_DN
#checked count of original df (df_DO_DN) and resulting df (df_DO_DN_fin) is the same

df_DO_DN <- combined_df_ref %>%
  filter(info_status_1 == "DO" & info_status_2 == "DN")

#this filter is used to identy cases that are DO vs. DN without further attributes; single phrase antecedent
df_DO_DN_a <- df_DO_DN %>%
  filter(ref_type == "phrase" & ref_type_2 == "new" &
           phrase_antecedent == "single_phrase" &
           related_object == "no" & related_object_2 == "no"
           )
#assign label
df_DO_DN_a$amb_reason_pre <- "DO (single) vs. DN"

#this filter is used to identy cases that are DO+bridging vs. DN; single phrase antecedent
df_DO_DN_a_1 <- df_DO_DN %>%
  filter(ref_type == "phrase" & ref_type_2 == "new" &
           phrase_antecedent == "single_phrase" &
           related_object == "yes" & related_object_2 == "no"
           )
#assign label
df_DO_DN_a_1$amb_reason_pre <- "DO+bridging (single) vs. DN"

#this filter is used to identy cases that are DO vs. DN+bridging single phrase antecedent
df_DO_DN_a_2 <- df_DO_DN %>%
  filter(ref_type == "phrase" & ref_type_2 == "new" &
           phrase_antecedent == "single_phrase" &
           related_object == "no" & related_object_2 == "yes"
           )
#assign label
df_DO_DN_a_2$amb_reason_pre <- "DO (single) vs. DN+bridging"

#this filter is used to identy cases that are DO (singel) + bridging vs. DN+bridging single phrase antecedent; same relation
df_DO_DN_a_3 <- df_DO_DN %>%
  filter(ref_type == "phrase" & ref_type_2 == "new" &
           phrase_antecedent == "single_phrase" &
           related_object == "yes" & related_object_2 == "yes" &
           related_rel == related_rel_2
           )
#assign label
df_DO_DN_a_3$amb_reason_pre <- "DO+bridging (single) vs. DN+bridging"

#this filter is used to identy cases that are DO (singel) + bridging vs. DN+bridging single phrase antecedent; different relation
df_DO_DN_a_4 <- df_DO_DN %>%
  filter(ref_type == "phrase" & ref_type_2 == "new" &
           phrase_antecedent == "single_phrase" &
           related_object == "yes" & related_object_2 == "yes" &
           related_rel != related_rel_2
           )
#assign label
df_DO_DN_a_4$amb_reason_pre <- "DO+bridging (single) vs. DN+bridging; ambiguous relation"

#this filter is used to identy cases that are DO vs. DN without further attributes; multiple phrase antecedents
df_DO_DN_b <- df_DO_DN %>%
  filter(ref_type == "phrase" & ref_type_2 == "new" &
           phrase_antecedent == "multiple_phrases" &
           related_object == "no" & related_object_2 == "no"
           )
#assign label
df_DO_DN_b$amb_reason_pre <- "DO (multiple) vs. DN"

#this filter is used to identy cases that are DO+bridging vs. DN ; multiple phrase antecedents --------------------------> no such cases
# df_DO_DN_b_1 <- df_DO_DN %>%
#   filter(ref_type == "phrase" & ref_type_2 == "new" &
#            phrase_antecedent == "multiple_phrases" &
#            related_object == "yes" & related_object_2 == "no"
#            )
# #assign label
# df_DO_DN_b_1$amb_reason_pre <- "DO+bridging (multiple) vs. DN"

#this filter is used to identy cases that are DO vs. DN+bridging ; multiple phrase antecedents
df_DO_DN_b_2 <- df_DO_DN %>%
  filter(ref_type == "phrase" & ref_type_2 == "new" &
           phrase_antecedent == "multiple_phrases" &
           related_object == "no" & related_object_2 == "yes"
           )
#assign label
df_DO_DN_b_2$amb_reason_pre <- "DO (multiple) vs. DN+bridging"

# #this filter is used to identy cases that are DO+bridging vs. DN+bridging ; multiple phrase antecedents, related relation is the same --------------------------> no such cases
# df_DO_DN_b_3 <- df_DO_DN %>%
#   filter(ref_type == "phrase" & ref_type_2 == "new" &
#            phrase_antecedent == "multiple_phrases" &
#            related_object == "yes" & related_object_2 == "yes" &
#            related_rel == related_rel_2
#            )
# #assign label
# df_DO_DN_b_3$amb_reason_pre <- "DO+bridging (multiple) vs. DN+bridging"

#this filter is used to identy cases that are DO+bridging vs. DN+bridging ; multiple phrase antecedents, related relation is not the same --------------------------> no such cases
# df_DO_DN_b_4 <- df_DO_DN %>%
#   filter(ref_type == "phrase" & ref_type_2 == "new" &
#            phrase_antecedent == "multiple_phrases" &
#            related_object == "yes" & related_object_2 == "yes" &
#            related_rel != related_rel_2
#            )
# #assign label
# df_DO_DN_b_4$amb_reason_pre <- "DO+bridging (multiple) vs. DN+bridging; ambiguous related relation"

#this filter is used to identy cases that are DO+deixis vs. DN without further attributes
df_DO_DN_c <- df_DO_DN %>%
  filter(ref_type == "segment" & ref_type_2 == "new" &
           related_object == "no" & related_object_2 == "no"
           )
#assign label
df_DO_DN_c$amb_reason_pre <- "deixis vs. DN"

#this filter is used to identy cases that are DO+deixis+bridging vs. DN without further attributes --------------------------> no such cases
# df_DO_DN_c_1 <- df_DO_DN %>%
#   filter(ref_type == "segment" & ref_type_2 == "new" &
#            related_object == "yes" & related_object_2 == "no"
#            )
# #assign label
# df_DO_DN_c_1$amb_reason_pre <- "deixis+bridging vs. DN"

#this filter is used to identy cases that are DO+deixis vs. DN+bidging without further attributes
df_DO_DN_c_2 <- df_DO_DN %>%
  filter(ref_type == "segment" & ref_type_2 == "new" &
           related_object == "no" & related_object_2 == "yes"
           )
#assign label
df_DO_DN_c_2$amb_reason_pre <- "deixis vs. DN+bridging"

#this filter is used to identy cases that are DO+deixis+bridging vs. DN+bidging without further attributes --------------------------> no such cases
# df_DO_DN_c_3 <- df_DO_DN %>%
#   filter(ref_type == "segment" & ref_type_2 == "new" &
#            related_object == "yes" & related_object_2 == "yes"
#            )
# #assign label
# df_DO_DN_c_3$amb_reason_pre <- "deixis+briidging vs. DN+bridging"


# recombine the dfs

df_DO_DN_fin <- rbind(df_DO_DN_a, df_DO_DN_a_1, df_DO_DN_a_2, df_DO_DN_a_3, df_DO_DN_a_4, df_DO_DN_b, df_DO_DN_b_2, df_DO_DN_c, df_DO_DN_c_2)

#print(df_DO_DN_fin)

```

##### DO vs. DO

```{r}
###DO-DO

df_DO_DO <- combined_df_ref %>%
  filter(info_status_1 == "DO" & info_status_2 == "DO")

# filter for identifying the right antecedent
df_DO_DO_a <- df_DO_DO %>%
  filter(ref_type == "phrase" & ref_type_2 == "phrase" &
           phrase_antecedent == "single_phrase" & phrase_antecedent_2 == "single_phrase" &
           single_phrase_antecedent != single_phrase_antecedent_2 &
           related_object == "no" & related_object_2 == "no" #thus, cannot have related_phrase
           )
#assign label
df_DO_DO_a$amb_reason_pre <- "identifying the right antecedent" #amb_ante_deixis and amb_bridging are always "no" in df_DO_DO_a

# filter for identifying the right antecedent + bridging in 1
df_DO_DO_b <- df_DO_DO %>%
  filter(ref_type == "phrase" & ref_type_2 == "phrase" &
           phrase_antecedent == "single_phrase" & phrase_antecedent_2 == "single_phrase" &
           single_phrase_antecedent != single_phrase_antecedent_2 &
           related_object == "yes" & related_object_2 == "no" 
           )
#assign label
df_DO_DO_b$amb_reason_pre <- "identifying the right antecedent + bridging in 1" 

# filter for assigning DO vs. DO; amb in generic and category added later
df_DO_DO_b1 <- df_DO_DO %>%
  filter(ref_type == "phrase" & ref_type_2 == "phrase" &
           phrase_antecedent == "single_phrase" & phrase_antecedent_2 == "single_phrase" &
           single_phrase_antecedent == single_phrase_antecedent_2 &
           related_object == "no" & related_object_2 == "no" 
           )
#assign label
df_DO_DO_b1$amb_reason_pre <- "DO vs. DO" ### -----------------------> one observation: GB mrcnf10g word_1137; NOT ambiguous

# filter for identifying the right antecedent + bridging in 2
df_DO_DO_c <- df_DO_DO %>%
  filter(ref_type == "phrase" & ref_type_2 == "phrase" &
           phrase_antecedent == "single_phrase" & phrase_antecedent_2 == "single_phrase" &
           single_phrase_antecedent != single_phrase_antecedent_2 &
           related_object == "no" & related_object_2 == "yes" 
           )
#assign label
df_DO_DO_c$amb_reason_pre <- "identifying the right antecedent + bridging in 2" 

# filter for identifying the right antecedent + bridging in both ---------------------------------------> no such cases
# df_DO_DO_c1 <- df_DO_DO %>%
#   filter(ref_type == "phrase" & ref_type_2 == "phrase" &
#            phrase_antecedent == "single_phrase" & phrase_antecedent_2 == "single_phrase" &
#            single_phrase_antecedent != single_phrase_antecedent_2 &
#            related_object == "yes" & related_object_2 == "yes"
#            )


# filter for identifying the right antecedent for bridging ---------------------------------------> no such cases
# df_DO_DO_c2 <- df_DO_DO %>%
#   filter(ref_type == "phrase" & ref_type_2 == "phrase" &
#            phrase_antecedent == "single_phrase" & phrase_antecedent_2 == "single_phrase" &
#            single_phrase_antecedent == single_phrase_antecedent_2 &
#            related_object == "yes" & related_object_2 == "yes"
#            )


# filter for antecedent vs. deixis
df_DO_DO_d <- df_DO_DO %>%
  filter(ref_type == "phrase" & ref_type_2 == "segment" &
           related_object == "no" & related_object_2 == "no")
#assign label
df_DO_DO_d$amb_reason_pre <- "antecedent vs. deixis"

# filter for antecedent+bridging vs. deixis ---------------------------------------> no such cases
# df_DO_DO_d_1 <- df_DO_DO %>%
#   filter(ref_type == "phrase" & ref_type_2 == "segment" &
#            related_object == "yes" & related_object_2 == "no")
# #assign label
# df_DO_DO_d_2$amb_reason_pre <- "antecedent+bridging vs. deixis"

# filter for antecedent vs. deixis+bridging ---------------------------------------> no such cases
# df_DO_DO_d_2 <- df_DO_DO %>%
#   filter(ref_type == "phrase" & ref_type_2 == "segment" &
#            related_object == "no" & related_object_2 == "yes")
# #assign label
# df_DO_DO_d_2$amb_reason_pre <- "antecedent vs. deixis+bridging"

# filter for antecedent+brdiging vs. deixis+bridging ---------------------------------------> no such cases
# df_DO_DO_d_3 <- df_DO_DO %>%
#   filter(ref_type == "phrase" & ref_type_2 == "segment" &
#            related_object == "no" & related_object_2 == "yes")
# #assign label
# df_DO_DO_d_3$amb_reason_pre <- "antecedent+bridging vs. deixis+bridging"


# filter for antecedent vs. deixis
df_DO_DO_e <- df_DO_DO %>%
  filter(ref_type == "segment" & ref_type_2 == "phrase" &
           related_object == "no" & related_object_2 == "no" 
         )
#assign label
df_DO_DO_e$amb_reason_pre <- "antecedent vs. deixis" 

# filter for antecedent vs. deixis+related_object
df_DO_DO_e1 <- df_DO_DO %>%
  filter(ref_type == "segment" & ref_type_2 == "phrase" &
           related_object == "yes" & related_object_2 == "no" 
         )
#assign label
df_DO_DO_e1$amb_reason_pre <- "antecedent vs. deixis+bridging" 

# filter for antecedent+related_object vs. deixis
df_DO_DO_e2 <- df_DO_DO %>%
  filter(ref_type == "segment" & ref_type_2 == "phrase" &
           related_object == "no" & related_object_2 == "yes" 
         )
#assign label
df_DO_DO_e2$amb_reason_pre <- "antecedent+bridging vs. deixis"

# filter for antecedent+related_object vs. deixis+related object ----------------------------------------> no such cases
# df_DO_DO_e3 <- df_DO_DO %>%
#   filter(ref_type == "segment" & ref_type_2 == "phrase" &
#            related_object == "yes" & related_object_2 == "yes" 
#          )
# #assign label
# df_DO_DO_e3$amb_reason_pre <- "antecedent+related_object vs. deixis"+related object

# filter for amb in deixis antecedent
df_DO_DO_f <- df_DO_DO %>%
  filter(ref_type == "segment" & ref_type_2 == "segment" &
           segment_antecedent != segment_antecedent_2 &
           related_object == "no" & related_object_2 == "no" 
         )
#assign label
df_DO_DO_f$amb_reason_pre <- "identifying the right segment for deixis"

# filter for amb in deixis antecedent + related object in 1
df_DO_DO_f1 <- df_DO_DO %>%
  filter(ref_type == "segment" & ref_type_2 == "segment" &
           segment_antecedent != segment_antecedent_2 &
           related_object == "yes" & related_object_2 == "no" 
         )

df_DO_DO_f1$amb_reason_pre <- "identifying the right segment for deixis + related_object in 1"

# filter for amb in deixis antecedent + related object in 2  ----------------------------------------> no such cases
# df_DO_DO_f2 <- df_DO_DO %>%
#   filter(ref_type == "segment" & ref_type_2 == "segment" &
#            segment_antecedent != segment_antecedent_2 &
#            related_object == "no" & related_object_2 == "yes" 
#          )
#assign label
# df_DO_DO_f2$amb_reason_pre <- "identifying the right segment for deixis + related_object in 2"

# filter for amb in deixis antecedent + related object in both ---------------------------------------> no such cases
# df_DO_DO_f3 <- df_DO_DO %>%
#   filter(ref_type == "segment" & ref_type_2 == "segment" &
#            segment_antecedent != segment_antecedent_2 &
#            related_object == "yes" & related_object_2 == "yes" 
#          )
# #assign label
# df_DO_DO_f3$amb_reason_pre <- "identifying the right segment for deixis + related_object in both"

# filter for single vs. multiple antecedents
df_DO_DO_g <- df_DO_DO %>%
  filter(ref_type == "phrase" & ref_type_2 == "phrase" &
           phrase_antecedent == "multiple_phrases" & phrase_antecedent_2 == "single_phrase" &
           related_object == "no" & related_object_2 == "no"
           )
#assign label
df_DO_DO_g$amb_reason_pre <- "single vs. multiple antecedents"

# filter for single vs. multiple antecedents
df_DO_DO_g4 <- df_DO_DO %>%
  filter(ref_type == "phrase" & ref_type_2 == "phrase" &
           phrase_antecedent == "single_phrase" & phrase_antecedent_2 == "multiple_phrases" &
           related_object == "no" & related_object_2 == "no"
           )
#assign label
df_DO_DO_g4$amb_reason_pre <- "single vs. multiple antecedents"

# filter for single vs. multiple antecedents and related object in 1 ---------------------------------------> no such cases
# df_DO_DO_g1 <- df_DO_DO %>%
#   filter(ref_type == "phrase" & ref_type_2 == "phrase" &
#            phrase_antecedent == "multiple_phrases" & phrase_antecedent_2 == "single_phrase" &
#            related_object == "yes" & related_object_2 == "no"
#            )
# #assign label
# df_DO_DO_g1$amb_reason_pre <- "single vs. multiple antecedents and related object in 1"

# filter for single vs. multiple antecedents and related object in 2 ---------------------------------------> no such cases
# df_DO_DO_g2 <- df_DO_DO %>%
#   filter(ref_type == "phrase" & ref_type_2 == "phrase" &
#            phrase_antecedent == "multiple_phrases" & phrase_antecedent_2 == "single_phrase" &
#            related_object == "no" & related_object_2 == "yes"
#            )
# #assign label
# df_DO_DO_g2$amb_reason_pre <- "single vs. multiple antecedents and related object in 2"

# filter for single vs. multiple antecedents and related object in both ---------------------------------------> no such cases
# df_DO_DO_g3 <- df_DO_DO %>%
#   filter(ref_type == "phrase" & ref_type_2 == "phrase" &
#            phrase_antecedent == "multiple_phrases" & phrase_antecedent_2 == "single_phrase" &
#            related_object == "yes" & related_object_2 == "yes"
#            )
# #assign label
# df_DO_DO_g3$amb_reason_pre <- "single vs. multiple antecedents and related object in both"

# filter for identifying the right antecedents
df_DO_DO_h <- df_DO_DO %>%
  filter(ref_type == "phrase" & ref_type_2 == "phrase" &
           phrase_antecedent == "multiple_phrases" & phrase_antecedent_2 == "multiple_phrases" &
           multiple_phrase_antecedents != multiple_phrase_antecedents_2 &
           related_object == "no" & related_object_2 == "no"
           )
#assign label
df_DO_DO_h$amb_reason_pre <- "identifying the right antecedents"

# filter for identifying the right antecedents + related_object in 1 ---------------------------------------> no such cases
# df_DO_DO_h1 <- df_DO_DO %>%
#   filter(ref_type == "phrase" & ref_type_2 == "phrase" &
#            phrase_antecedent == "multiple_phrases" & phrase_antecedent_2 == "multiple_phrases" &
#            multiple_phrase_antecedents != multiple_phrase_antecedents_2 &
#            related_object == "yes" & related_object_2 == "no"
#            )
# #assign label
# df_DO_DO_h1$amb_reason_pre <- "identifying the right antecedents + related_object in 1"

# filter for identifying the right antecedents + related_object in 2 ---------------------------------------> no such cases
# df_DO_DO_h2 <- df_DO_DO %>%
#   filter(ref_type == "phrase" & ref_type_2 == "phrase" &
#            phrase_antecedent == "multiple_phrases" & phrase_antecedent_2 == "multiple_phrases" &
#            multiple_phrase_antecedents != multiple_phrase_antecedents_2 &
#            related_object == "no" & related_object_2 == "yes"
#            )
# #assign label
# df_DO_DO_h2$amb_reason_pre <- "identifying the right antecedents + related_object in 2"

# filter for identifying the right antecedents + related_object in both ---------------------------------------> no such cases
# df_DO_DO_h3 <- df_DO_DO %>%
#   filter(ref_type == "phrase" & ref_type_2 == "phrase" &
#            phrase_antecedent == "multiple_phrases" & phrase_antecedent_2 == "multiple_phrases" &
#            multiple_phrase_antecedents == multiple_phrase_antecedents_2 &
#            related_object == "yes" & related_object_2 == "yes"
#            )
# #assign label
# df_DO_DO_h3$amb_reason_pre <- "identifying the right antecedents + related_object in both"


### combinig the dfs; df_DO_DO_b1 is not included, since one observation that is NOT ambiguous DO vs. DO
df_DO_DO_fin <- rbind(df_DO_DO_a, df_DO_DO_b, df_DO_DO_c, df_DO_DO_d, df_DO_DO_e, df_DO_DO_e1, df_DO_DO_e2, df_DO_DO_f, df_DO_DO_f1, df_DO_DO_g, df_DO_DO_g4, df_DO_DO_h)
```

##### Recombine all Data Frames that Contain Referring Markables

```{r}
#recombine the categorised dfs of referring markables
df_categorised_referring <- rbind(df_DN_DN_fin, df_DO_DO_fin, df_DO_DN_fin, df_DN_DO_fin, df_ref_undef_ref)
print(df_categorised_referring)

```

##### Recombine all categorised Data Frames

```{r}
#recombine data frames
df_categorised_1 <- rbind(df_categorised_referring, df_non_ref)

print(df_categorised_1)
```

##### Checking Double Appearing Markables

```{r}
#check if a markable does not appear twice or more in the df
double_markables <- df_categorised_1 %>%
  group_by(file_name) %>% # group by the file the markable appears in
  filter(duplicated(id) | duplicated(id, fromLast = TRUE)) # use function to check for duplicate ids

print(double_markables) #if this data frame is empty, there are no markables that appear twice in the original df
```

##### Ambiguity in Generic and/or Category

```{r}
#create final data frame
df_categorised_fin <- df_categorised_1

#get rid of the superfluous semicolons 
concatenate_with_condition <- function(row, columns, separator = "; ", include_empty = FALSE) {
  selected_row <- row[columns]
  
  if (!include_empty) {
    selected_row <- selected_row[selected_row != ""]
  }
  
  result <- paste(selected_row, collapse = separator)
  return(result)
}

#specify columns to use
columns_to_use <- c("amb_reason_pre", "amb_reason_gen", "amb_reason_cat")

#concatenate and put in amb_reason_fin
df_categorised_fin$amb_reason_fin <- apply(df_categorised_fin[, columns_to_use], 1, concatenate_with_condition, 
                     columns = columns_to_use, separator = "; ", include_empty = FALSE)

```

#### Full List of Labels and Count

```{r, echo=FALSE}
print("List of labels and their count in ARRAU and PD:")
table(df_categorised_fin$amb_reason_fin)

print(df_categorised_fin)
print(combined_df_amb)
```

